---
title: "MINA Manual"
author: "Rui Guan"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MINA Manual}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
# Overview

The package is used for microbial profiling data analysis, including community
similarity / dissimilarity / distance calculation and network comparison.

# About this vignette

## Typesetting Legend <a id="sec:typeset-legend"></a>

- **bold** - Bold is used for emphasis.
- *italics* - Italics are used for package names, and special words, phrases.
- `code font` - The font for code, usually courrier-like, but depends on the
  theme.
- `myFun()` - Code font word with `()` attached at the right-end, is a functioni
  name.
- [Hyperlink](#sec:typeset-legend) - Hyperlinks are
clickable text that will jump to sections and external pages.

## Import data

The first step is to import quantitative table of microbial community profiling
data (e.g. OTU table) and corresponding descriptive file (e.g. design file) into
the object.
To load data, use the `new()` for all avaliable file or `addTab()` and
`addDes()` to import them seperately.

For the quantitative file (e.g. OTU table), rows are compositions of the
community and columns are samples. The corresponding descriptive file(e.g.
design file) should
have the same samples as quantitative table. The first column of descriptive
file is "**SampleID**" which contains IDs of samples.

#### Example

Import files together:

```{r, eval=FALSE}
m <- new(mina, Tab = "otu_table.txt", desTab = "design.txt")
```
Import files seperately:
```{r, eval=FALSE}
m1 <- new(mina)
m1 <- addTab(m1, "otu_table.txt")
m1 <- addDes(m1, "design.txt")
```
### Normalize quantitative table


## Supported Analysis


## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
